---
title: "01_EDA"
author: "Suvendu K. Pati"
date: "03/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(fpp3)
library(dplyr)
library(tsibble)
library(lubridate)
library(ggplot2)
```

## Importing datasets
```{r}
calendar <- read.csv("calendar_afcs2021.csv")
train <- read.csv("sales_train_validation_afcs2021.csv")
test <- read.csv("sales_test_validation_afcs2021.csv")
sale_price <- read.csv("sell_prices_afcs2021.csv")
sample <- read.csv("sample_submission_afcs2021.csv")
```

## Checking missingness
```{r}
sum(is.na(train))
sum(is.na(test))
```

## Number of products
```{r}
length(unique(train$id))
```

## Data transformation
### Transpose function
```{r}
transpose <- function(df) {
  tdf <- t(df)
  col <- df[,1]
  tdf <- as.data.frame(tdf[-1,]) 
  colnames(tdf) <- col  
  tdf <- as.data.frame(tdf)
}
```

Transposing both train and test data to obtain data in desired format
```{r}
trainT <- transpose(train)
trainT <- tibble::rownames_to_column(trainT, "d")

testT <- transpose(test)
testT <- tibble::rownames_to_column(testT, "d")
```

### Data type casting
#### Changing character to numeric columns
```{r}
nm <- colnames(trainT)
nm <- nm[-1]
for(cols in nm){
  trainT[,cols] <- as.numeric(trainT[,cols])
}
```

### New column for total number of products sold
```{r}
trainT$total <- rowSums(trainT[,nm])
```

## Merging calendar and training set
```{r}
train_new <- merge(calendar, trainT, by = "d")
sum(is.na(train_new))
```

### changing character dates to date type
```{r}
train_new$date <- mdy(train_new$date)
```

### Converting to tsibble object type
```{r}
train_ts <- train_new %>%
  as_tsibble(
    index = date
    )
```

# Number of events, monthwise
```{r}
train_new %>%
  select(year, month, event_name_1, event_type_1, event_name_2, event_type_2) %>%
  mutate(
    event = if_else((event_name_1 != "" | event_type_1 != "" | event_name_2 != "" | event_type_2 != ""), 1, 0)
  ) %>%
  group_by(year, month) %>%
  summarise(event = sum(event)) %>%
  filter(year==2012) %>%
  ggplot(aes(x=month, y=event, group=year)) +
  geom_line(aes(color=year)) +
  scale_x_continuous(name = "Months", breaks = c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  labs(y="Number of events")
  
```

## Plotting
### Total product sold
```{r}
train_ts %>% 
  filter(year(date)==2012) %>%
  autoplot(total) +
  labs(x="Year", y="Total number of products sold")
```

```{r}
train_ts %>% gg_season(total)
```

```{r}
wkdaysales <- train_new %>% 
  select(weekday, month, year, total) %>%
  group_by(year, month, weekday) %>%
  summarise(total = sum(total)/1e05)

ggplot(wkdaysales, aes(x=weekday, y=total)) + 
  geom_bar(stat = "identity") +
  labs(y="total sales (x 100000)")

ggplot(wkdaysales, aes(x=month, y=total)) + 
  geom_bar(stat = "identity") +
  scale_x_continuous(name = "Months", breaks = c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  labs(y="total sales (x 100000)")

ggplot(wkdaysales, aes(x=year, y=total)) + 
  geom_bar(stat = "identity") +
  labs(y="total sales (x 100000)")
```

```{r}
wkdaysales <- train_new %>% 
  select(weekday, month, year, total) %>%
  group_by(year, month, weekday) %>%
  summarise(total = sum(total)/1e05)
```


### Plot for individual product
```{r}
train_ts %>% 
  autoplot(FOODS_3_001_TX_3_validation) +
  labs(x="Year", y="Total number of FOODS_3_001_TX_3_validation sold")

train_ts %>% 
  autoplot(FOODS_3_002_TX_3_validation) +
  labs(x="Year", y="Total number of FOODS_3_002_TX_3_validation sold")

train_ts %>% 
  autoplot(FOODS_3_003_TX_3_validation) +
  labs(x="Year", y="Total number of FOODS_3_003_TX_3_validation sold")

train_ts %>% 
  autoplot(FOODS_3_007_TX_3_validation) +
  labs(x="Year", y="Total number of FOODS_3_007_TX_3_validation sold")
```

## Decomposition of the total number of products sold
```{r}
dcmp <- train_ts %>%
  model(stl = STL(total))

components(dcmp) %>% autoplot()
```

## Decomposition for specific product
```{r}
dcmp <- train_ts %>%
  model(stl = STL(FOODS_3_002_TX_3_validation))

components(dcmp) %>% autoplot()
```

```{r}
evntwise_sale <- train_new %>%
  select(year, event_type_1, total) %>%
  #if_else(event_type_1 != "", event_type_1, "None") %>%
  group_by(year, event_type_1) %>%
  summarise(total=mean(total))

evntwise_sale$event_type_1 <- ifelse(evntwise_sale$event_type_1=="","None",evntwise_sale$event_type_1)

evntwise_sale %>%
  ggplot(aes(x=year, y=total, fill=factor(event_type_1))) +
  geom_line() +
  facet_grid(vars(event_type_1), scales = "free_y") +
  labs(y="Mean sales")
```

```{r}
evntwise_sale <- train_new %>%
  select(year, event_type_1, total) %>%
  group_by(year, event_type_1) %>%
  summarise(total=mean(total))

evntwise_sale$event_type_1 <- ifelse(evntwise_sale$event_type_1=="","None",evntwise_sale$event_type_1)

evntwise_sale %>%
  ggplot(aes(x=event_type_1, y=total)) +
  geom_bar(stat = "identity") +
  labs(x="Holiday types", y="Mean sales")
```

```{r}
evntwise_sale %>%
  pivot_wider(values_from=total, names_from=event_type_1) %>%
  GGally::ggpairs(2:6)
```

### Modelling
```{r}
fit <- train_ts %>% 
  model(
    trend_model = TSLM(FOODS_3_001_TX_3_validation ~ trend()),
    MEAN = MEAN(FOODS_3_001_TX_3_validation),
    NAIVE = NAIVE(FOODS_3_001_TX_3_validation),
    SNAIVE = SNAIVE(FOODS_3_001_TX_3_validation),
    Drift = RW(FOODS_3_001_TX_3_validation ~ drift())
    )

fit %>% accuracy()
```

```{r}
fit %>% gg_tsresiduals()
```

